---
title: "UPC & IDBR Business Premises"
author: "Fushu Beauthier"
date: "`r Sys.Date()`"
output: html_document
---
[Descriptive text]

N.B. For each file, the directory/file path and file names in the code below need to be modified according to where you have stored each file and how you saved it when you downloaded it.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

**Loading the necessary packages**
```{r}
library(ggpubr)
library(tidyverse)
library(qacEDA)
library(gstat)
```

#### Load the UPC estimated business premises 20XX, pulled from Snowflake table
```{r}

```

#### Load the IDBR business premises, last updated 2019 (as of 2023)
```{r}
# change directory/file path as required #
# idbr <- read_csv("data/demos/")
# 
# idbr <- idbr %>%
#   rename(MSOA = 1, Total = 2) %>%
#   filter(!row_number() %in% c(1:7))
# 
# # edit MSOA to codes only
# idbr$MSOA <- sapply(strsplit(idbr$MSOA, ":"), "[", 2)
# idbr$MSOA <- sapply(strsplit(idbr$MSOA, " "), "[", 1)
# head(idbr$MSOA)
# idbr$Total <- as.integer(idbr$Total)
# summary(idbr)

```


#### Load the IDBR local units, last updated 2019 (as of 2023)
```{r}

```

